Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks

### 2022-05-10 rmd/manuscript.Rmd
edit: author info

**Author:** Justin P.F. Pomeranz^1,2,\*^, Jacques C. Finlay^3^, Akira Terui^1,\*^

^1^ Department of Biology, University of North Carolina at Greensboro; ^2^ Department of Environmental Sciences, Colorado Mesa University; ^3^ Department of Ecology, Evolution and Behavior, University of Minnesota

**\*Corresponding Author:** JPF Pomeranz, jfpomeranz\@gmail.com; A Terui, hanabi0111\@gmail.com

@@

Under each scenario, we ran `r max(sim_main_result$n_rep)` simulation replicates with `r n_sim` timesteps of trophic dynamics (including `r n_sim - unique(sim_main_result$n_timestep)` steps of warmup and burn-in). During the warmup period (`r unique(sim_main_result$n_warmup)` time steps), we introduced propagules into habitat patches every 10 time steps to allow species to establish populations. Propagules are random draws from $Poisson(0.5)$. Subsequently, we ran a burn-in period with no propagules for `r unique(sim_main_result$n_burnin)` time steps to remove the effects of initial conditions. We obtained the mean FCL across all habitat patches for the last `r unique(sim_main_result$n_timestep)` time steps. Note that simulation replicates are differ in ecosystem size ($N_{p}\sim Unif(10,150)$) and ecosystem complexity ($P_{b} \sim Unif(0.01, 0.99)$) so we can estimate their effects on FCL.

@@

# **Acknowledgements**

We thank C Perovich for thoughtful discussions in the early stages of this project. This material is based upon work supported by the National Science Foundation through the Division of Environmental Biology (DEB 2015634).

# **Author contributions**

JP developed the theoretical model of trophic dynamics with input from AT. JP and AT ran the simulations, and made the figures. JP wrote the first draft, while AT and JF contributed significantly to the final version.

# **Data availability**

No data were used or produced in this research. R functions for simulations are provided as the R package *mcbrnet* (available at <https://github.com/aterui/mcbrnet>). Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.

>>>>

@@

Under each scenario, we ran `r max(sim_main_result$n_rep)` simulation replicates with `r n_sim` timesteps of trophic dynamics (including `r n_sim - unique(sim_main_result$n_timestep)` steps of warmup and burn-in). During the warmup period (`r unique(sim_main_result$n_warmup)` time steps), we introduced propagules into habitat patches every 10 time steps to allow species to establish populations. Propagules are random draws from $Poisson(0.5)$. Subsequently, we ran a burn-in period with no propagules for `r unique(sim_main_result$n_burnin)` time steps to remove the effects of initial conditions. We obtained the mean FCL across all habitat patches for the last `r unique(sim_main_result$n_timestep)` time steps. Note that simulation replicates are differ in ecosystem size ($N_{p}\sim Unif(10,150)$) and ecosystem complexity ($P_{b} \sim Unif(0.01, 0.99)$) so we can estimate their effects on FCL. All analysis was performed in R 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021]

@@
